### 2021-12-07 man/floa_doc.Rmd
Move files to man folder

---
title: "Documentation/overview Functional Limits of Agreement"
author:
  - name: Daniel Koska
    email: daniel.koska@hsw.tu-chemnitz.de
    affiliation: Chemnitz University of Technology, Research Methodology and Data Analysis in Biomechanics
    footnote: Corresponding Author
date: "`r format(Sys.time(), '%d %B, %Y')`"
# output: github_document
output:
  html_document:
    keep_md: yes
    theme: united
    toc: yes
    toc_float: yes
---

## Zusammenfassung

#### Übergeordnete Fragestellung

"What is the coverage performance of functional prediction bands when analyzing the agreement between two sets of biomechanical curves?"

Frage ergab sich ursprünglich im Rahmen der Validierung einer hauseigenen IMU gegenüber einem kamerabasierten Messsystem (Vicon).

Standardvorgehen: Bland & Altman Plots (diskrete Parameter)

Erweiterung für funktionale Daten: Roislien et al. (2012) Functional Limits of Agreement

#### Functional Limits of Agreement nach Roislien et al.

+ Behandlung von Zeitreihen als funktionale Objekte (Ramsay & Silverman)

+ Auswahl einer Zeitreihe je Proband (bei n=7! im Beispiel) und Schätzung der LoA durch Bootstrapping

#### Koska & Maiwald (2020, submitted to Gait & Posture) ... Vergleich (coverage) 3er Verfahren:

+ Pointwise (continuous) LoA (FLoA~Point~)

+ Functional Limits of Agreement nach Roislien et al. (FLoA~2SD~)

+ Functional limits of agreement using a randomized cluster bootstrap (FLoA~RCB~, Koska & Maiwald)

Illustriert wurden die Verfahren am Beispiel eines Datensatzes zur Validierung einer IMU (MoCap vs. IMU)

#### Methodikbeschreibung FLoA~RCB~ (kopiert aus dem Paper)

- The agreement between the two measurement systems was analyzed based on time-normalized difference curves

- Initially, time-discrete curves 𝑑⃗(𝑡) were converted to functional objects 𝑥(𝑡) using a linear combination of 𝐾 (Fourier series) functions with weight coefficients 𝑐
  + 𝐾 was set relatively high (𝐾=50) to avoid smoothing artifacts when fitting the function. The quality of the fit was checked visually using plots.

This model includes intra-individual variance from repeated measurements. To account for the cluster structure generated by similar curves of the same subject (Fig. 1), we implemented the randomized cluster bootstrap described in Davison & Hinkley [20]. In the first stage, 𝑛 clusters were drawn with replacement from the original sample. A single cluster corresponds to all curves of a subject. In the second stage, observations within the newly generated sample were randomly permuted (drawing without replacement). The process was repeated 𝐵 times and mean curves were determined during every iteration.

From the resulting empirical distribution of mean curves 𝑋𝐵∗𝑛∗, pointwise quantiles (Q2.5, Q97.5) were determined as upper and lower interval limits: 𝐹𝐿𝑜𝐴𝑅𝐶𝐵= ± 𝑄𝛼/2(𝑋𝑖 𝐵∗𝑛 ∗(𝑡))

## Flowchart FLoA~RCB~

#### Eingereichte Version (problematisch)

```{r echo = FALSE, warning = FALSE, message = FALSE}
library(DiagrammeR)

# (Nested) Randomized Cluster Bootstrap
#
# In the first stage (draw_clusters()), all curves of a single (random) subject
# are drawn. From this set, the (functional) mean is calculated. The process is
# repeated n.boot times (n.boot = number of bootstrap iterations)

grViz("digraph flowchart {

      # node definitions with substituted label text
      node [fontname = Helvetica, shape = rectangle]

      tab1 [label = '@@1']
      tab2 [label = '@@2']
      tab3 [label = '@@3']
      tab4 [label = '@@4']
      tab5 [label = '@@5']
      tab6 [label = '@@6']
      tab7 [label = '@@7']
      tab8 [label = '@@8']
      tab9 [label = '@@9']
      tab10 [label = '@@10']

      # edge definitions with the node IDs
      tab1 -> tab2
      tab2 -> tab3
      tab3 -> tab4
      tab3 -> tab5
      tab3 -> tab6
      tab4 -> tab7
      tab7 -> tab8
      tab8 -> tab9
      tab9 -> tab10
      }

      [1]: 'Time normalize curves of IMU - MC'
      [2]: 'Create difference curves IMU - MC'
      [3]: 'Convert difference curves to function data objects using Fourier series'
      [4]: 'floa_rcb '
      [5]: 'FLOAboot_2SD '
      [6]: 'floa_point'
      [7]: paste0('First stage', ':\\n ', 'draw_clusters', '\\n ', 'All strides from 11 subjects (randomly permuted)', '\\n ', 'are drawn with replacement')
      [8]: 'Functional mean for each cluster'
      [9]: paste0('Second stage', ':\\n ', 'Repeat the first stage 1000 times')
      [10]: 'Quantile'
      ")

```

Asymmetrische Limits of Agreement aus einem hierarchischen (zweistufigen) Bootstrap in Anlehnung an Davison & Hinkley (1997)

#### Vorschlag für verbesserte Variante

## Reviewerkommentare (sortiert nach Wichtigkeit)

#### Vorab etwas Positivität

+ Reviewer #1: I would like to congratulate the authors for a complex, well written and meaningful manuscript. [...] I have only minor comments: ...

+ Reviewer #2: The general idea behind the manuscript, to evaluate agreement between biomechanical curves and to construct prediction bands are highly relevant and interesting.

+ Reviewer #3: The idea that the authors are presenting is quite interesting, but theanalyis and evaluation of the various mathods contains come serious errors.

#### Major comments

+ Evaluierung anhand weiterer (simulierter) Datensätze

+ In looking at limits of agreement, the authors are considering the extent of within pair variation. However, in looking at the point wise limits of agreement, these are defined to incorporate a between pair standard deviation. Even if it was appropriate to utilise between pair variability, the authors should note that it is variances that are additive and not SDs. Perhaps the authors lack of understanding comes from their *erroneous assumption* (iii).

*iii Measurement uncertainty can be estimated adequately by bootstrapping the error variance from a sample that consists of a single curve per subject.*

*If no repeated measurements are included in the sample, an essential error component (intra-individual variance) is ignored. The bootstrap – like any other statistical procedure – cannot solve this problem, as it can only estimate proportions of variance that are contained in the sample.*

Die erwähnte assumption bezieht sich auf eine Grundannahme die im Roislien Paper getroffen wurde und in unserem Paper kritisch aufgegriffen wurde.

+ [...] The authors are producing the arithmetic mean curves, the variability of which will reflect variation between means. It is the variability of individual values that is relevant, and the limits will be too narrow by a factor of approximately the square root of 11.

+ As I understand the description of the paper, the coverage is estimated for the curves that are used to estimate the prediction bands

+ Detaillierte Beschreibung der Methode erforderlich

+ Mathematische Notation

#### Fragwürdige Einschätzungen seitens der Reviewer

+ Fourier Series sind overfitted: "The effect of overfitting will be that the fitted curves wll reflect to a large extent the measurement errors associated with each individual point."

### Mögliche Lösungsansätze

+ Herausnehmen der punktweisen LoA: Die Idee funktionaler Methode könnte man als etabliert betrachten. Beim Lesen der Reviewerkommentare drängt sich auch der Verdacht auf, dass die relativ hohe Komplexität des Papers sich negativ auf das Verständnis (der Reviewer) auswirkt.

... siehe dieser Kommentar der Reviewer: Although the pointwise approach ignores the temporal dependence, it is not an issue when it comes to estimation. For inference, on the other hand, it should be taken into account

+ Die FLoA~2SD~ (Roislien et al.) würde ich hingegen erstmal drin lassen, da a. unsere Methode darauf aufbaut und b. die Vorteile unserer Methode so besser veranschaulicht werden können. Zudem kann ich mir vorstellen, dass die Forderung nach einer Vergleichsmethode beim Einreichen eines neuen Papers zwangsläufig irgendwann kommt.

+ Datensätze mit ähnlichen Eigenschaften wie typische biomechanische Signale simulieren (z. B. Beschleunigungssignale, EMG, EEG)

... evtl. kommen die (vermuteten) Vorteile der FLoA~RCB~ bei Signalen mit höherem Rauschanteil besser zum Tragen.

## Sonstiges

+ FLoA~RCB~ als R package

+ Github repository: https://github.com/koda86/floa